---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vocabular2

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- [![CRAN status](https://www.r-pkg.org/badges/version/vocabr)](https://CRAN.R-project.org/package=vocabr) -->
<!-- badges: end -->

The goal of vocabular2 is to compare vocabularies on a set of metrics. 
There's currently no clear development path for the package. It may become usable in the future, but for now it's not adviced to use the code for your projects.

## Installation

You can install the development version with:

``` r
devtools::install_github("ludvigolsen/vocabular2")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(vocabular2)
## basic example code
```


